CSE 517: Project Proposal

### 2019-03-18 cse517/proposal/main.tex
Develop paper (#14)
* First draft of project proposal.

* Ready to submit.

* Switching to emnlp template.

* Moving proposal content to paper.

* Adding figures.

* Procrastinating by adding random text to paper.

* Starting report.

* Updating report

\begin{document}
\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

@@

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts.
We seek to improve NLI models by adding syntactic
features into existing models. The well-known decomposable attention (DA)
neural network \citep{Parikh2016-em} obtained state of the art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.

We intend to extend this
model and evaluate it on SNLI, MultiNLI \citep{Williams2017-uh}, the SNLI and
MultiNLI hard splits \citep{Gururangan2018-lj}, and smaller domain specific
datasets like Scitail \citep{Khot2018-th}.

@@

In previous work, we created the syntail-v1 model whose architecture is
shown in figure \ref{figure:v1}.
We have obtained mixed results as shown in table \ref{table:v1-accuracies} when
naively incorporating syntactic information
into the DA model.

Wihthout ELMo, syntail-v1 improves an impressive 8.4\% in test accuracy over
the baseline decomposable attention model on the SciTail dataset. However,
once ELMo is added to the model, almost the exact happens. Adding syntactic
information decreases test accuracy by 6.7\%.

\begin{table}[h]
  \resizebox{0.45\textwidth}{!}{
  \begin{tabular}{c c c c}
    \toprule
    Model & Embedding & Train Acc. & Test Acc. \\
    \midrule
    DA & GLoVe 6b 300d & 89.4\% & 70.1\% \\
    \textbf{Syntail-v1} & \textbf{GLoVe 6b 300d} & \textbf{92.1\%} & \textbf{78.5\%} \\
    \textbf{DA} & \textbf{ELMo} & \textbf{98.4\%} & \textbf{78.0\%} \\
    Syntail-v1 & ELMo & 88.8\% & 71.3\% \\
    \bottomrule
  \end{tabular}}
  \caption{Syntail-v1 and DA model accuracies on the SciTail dataset.}
\label{table:v1-accuracies}
\end{table}

@@

We plan to use state-of-the-art constituency \citep{Stern2017-co} and
dependency \citep{Dozat2016-gs} parsers to investigate if incorporating
syntactic features improves the performance of the decomposable attention
network. Additionally, we hope to understand how syntactic features interact
with contextual
word embeddings like ELMo \citep{Peters2018-fz} and BERT \citep{Devlin2018-qc}
in a task like NLI.

We will first investigate different ways of incorporating syntactic
information. \citet{Pang2018-syntail} use a constituency parser to obtain an
encoding of the premise and hypothesis that is used when characterizing their
relationship. We believe that incorporating syntactic features earlier in the
pipeline, especially before the premise and hypothesis are attended over one
another, should improve performance.

>>>>

\begin{document}
\maketitle

@@

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts.  We seek to improve models which
perform NLI by adding syntactic features into existing models. In particular,
we will primarily investigate the well-known decomposable attention (DA) neural
network \citep{Parikh2016-em} which obtained state of the art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.

We intend to extend the DA model and evaluate new models on SNLI, MultiNLI
\citep{Williams2017-uh}, the SNLI and MultiNLI hard splits
\citep{Gururangan2018-lj}, and smaller, domain specific datasets like Scitail
\citep{Khot2018-th}.

In general, most NLI tasks are formulated as characterizing the relationship
between a pair of sequences---a premise and a hypothesis. An NLI model should
predict whether the hyothesis is entailed by the premise, contradicts the
premise, or is neutral to the premise.

@@

In previous work, we created the syntail-v1 model whose architecture is shown
in figure~\ref{figure:v1}.  We have obtained mixed results as shown in
table~\ref{table:v1-accuracies} when naively incorporating syntactic
information into the DA model.

Without ELMo, syntail-v1 improves an impressive 8.4\% in test accuracy over the
baseline decomposable attention model on the SciTail dataset. However, once
ELMo is added to the model, almost the exact opposite occurs. Adding syntactic
information decreases test accuracy by 6.7\% with ELMo.

\begin{table}[h]
  \resizebox{0.45\textwidth}{!}{
  \begin{tabular}{c c c c}
    \toprule
    Model & Embedding & Train Acc. & Test Acc. \\
    \midrule
    DA & GLoVe 6b 300d & 89.4\% & 70.1\% \\
    \textbf{Syntail-v1} & \textbf{GLoVe 6b 300d} & \textbf{92.1\%}
                        & \textbf{78.5\%} \\
    DA & ELMo & 98.4\% & 78.0\% \\
    \textbf{Syntail-v1} & \textbf{ELMo} & \textbf{88.8\%} & \textbf{71.3\%} \\
    \bottomrule
  \end{tabular}}
  \caption{Syntail-v1 and DA model accuracies on the SciTail dataset.}
\label{table:v1-accuracies}
\end{table}

@@

We plan to use state-of-the-art constituency \citep{Stern2017-co} and
dependency \citep{Dozat2016-gs} parsers to investigate if incorporating
syntactic features improves the performance of the DA model. Additionally, we
hope to understand how syntactic features interact with contextual word
embeddings like ELMo \citep{Peters2018-fz} and BERT \citep{Devlin2018-qc} in a
task like NLI. It is unclear from previous work whether contextual word
embeddings benefit from additional syntactic information or if something like
ELMo can already adequately capture this information.

We will first investigate different ways of incorporating syntactic
information. \citet{Pang2018-syntail} use a constituency parser to obtain an
encoding of the premise and hypothesis that is in the final fully-connected
layer of the DA model. We believe that incorporating syntactic features earlier
in the pipeline, especially before the premise and hypothesis are attended over
one another, should improve performance.
